---
title: "classmate_analysis"
output: html_document
date: "2023-01-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```

# Youth Classmates Network Analysis

```{r load packages}
library(dplyr)
if(!require(statnet)) install.packages("statnet",repos = "http://cran.us.r-project.org")
library("statnet")
library(GGally)
library(tidyverse)
```

#### 65% of all 80 schools in our subsample examined 2 classes

```{r count classes by school}
# total number of school -- 80
w1_yc <- haven::read_dta("Data-Stata/our_data/w1_yc.dta")
w1_yc %>% 
  group_by(schoolidCUF) %>% 
  summarise(n_class = length(unique(classidCUF))) %>% 
  group_by(n_class) %>% 
  summarise(count_by_num_class=n(),
            percent = count_by_num_class/80)
```

```{r count students by school}
w1_yc %>% 
  group_by(schoolidCUF) %>% 
  summarise(total_obs = n()) %>% # NUM of participants from each school
  arrange(desc(total_obs)) %>% 
  top_n(5)
```

```{r use school3928--highest # of participated students as example}
school3928 <- w1_yc %>% 
  filter(schoolidCUF == 3928)

# exclude useless variables
school3928 <- school3928 %>% 
  select(-schtype_geRV, -y1_intdat_ycRV, -y1_vers_yc)
#colnames(school3928)
```

[Tutorial on labelled/unlabelled object in R](https://cran.r-project.org/web/packages/labelled/vignettes/intro_labelled.html)

```{r data preprocess for school3928 network}
# data only includes friendship measures
network3928 <- school3928 %>% 
  select(youthidCUF, 4:210)

# code out missing values
network3928_clean <- network3928 %>% 
  labelled::to_factor() %>%  # changed labelled/haven_labelled object to factor
  lapply(function(x) as.numeric(as.character(x))) %>% # coerce all vars to be numeric
  data.frame()
```

Relevant tutorial on networks in R [HERE](https://rstudio-pubs-static.s3.amazonaws.com/470420_91f1d47c90c54b879aa08ce93fa66730.html)

Storing a network using an adjacency matrix also helps storing network attributes (by using the rownames of the adjacency matrix and assigning those rownames the corresponding attributes - we'll see that later).

Following the workflow described above, we are going to create an empty matrix with the unique student IDs as the row and column names so that we can later populate them with the information from the edgelist.

```{r construct adjacency matrix}
students <- unique(network3928_clean$youthidCUF)
mat <- matrix(0, nrow=length(students), ncol = length(students))
colnames(mat) <- as.character(students)
rownames(mat) <- as.character(students)
# A glimpse of what this look lik
mat[1:4,1:4]
```

We are now going to use a loop to fill the matrix: 1 whenever there is a friendship tie in students, 0 otherwise.

```{r fill adjacency matrix}
i <- 1
for (i in 1:nrow(network3928_clean)) {
  row.index <- which(students == network3928_clean[i,1])
  col.index <- which(students == network3928_clean[i,3]) # col 3 is best friends 1
  mat[row.index, col.index] <- 1
}
mat[1:4,]
```

### Friendship Network at School 3928 --\> All 5 best friends

Using `pivot_longer()` functions in `tidyverse` to reshape our data from wide to long ([Relevant Tutorial here](https://tavareshugo.github.io/r-intro-tidyverse-gapminder/09-reshaping/index.html))

```{r reshape data}
bfs3928 <- network3928_clean %>% 
  select(youthidCUF, y1_bfs_1:y1_bfs_5) %>% 
  pivot_longer(cols = y1_bfs_1:y1_bfs_5,
               names_to = "best_friends",
               values_to = "friendid")

bfs3928 <- bfs3928 %>% 
  select(youthidCUF, friendid) %>% 
  filter(!is.na(friendid)) %>% 
  mutate(youthidCUF = as.character(youthidCUF),
         friendid = as.character(friendid))

bfs_nw <- network(bfs3928, matrix.type = "edgelist")
#Or we can do it by using the adjacency matrix:
#nw2 <- network(mat, directed = TRUE)
ggnet2(bfs_nw,
       label = TRUE,
       label.size = 3,
       arrow.size = 3,
       arrow.gap = .03)
```

# count th number of mutual friends

# social ties & instrumental ties

# All Schools

```{r data preprocess for all schools}
# data only includes friendship measures
w1_pos <- w1_yc %>% 
  select(youthidCUF, 7:210) %>% 
  mutate(y1_bfs_0 = as.numeric(as.character(y1_bfs_0)))

# code out missing values
w1_pos <- w1_pos %>% 
  #labelled::to_factor() %>%  # changed labelled/haven_labelled object to factor
  lapply(function(x) as.numeric(as.character(x))) %>% # coerce all vars to be numeric
  data.frame() 

w1_pos[w1_pos < 0] <- NA

# exclude the 4 students did not respond to best friends questions at all
descr::freq(w1_yc$y1_bfs_0, plot=FALSE)
descr::freq(w1_pos$y1_bfs_0, plot=FALSE)
#na_students <- c(52280219,42110503,31080205,29840209)
#w1_pos <- w1_pos %>% 
#  filter(!youthidCUF %in% na_students)
```

```{r reshape data to be long format}
w1_pos_long <- w1_pos %>% 
  filter(!is.na(y1_bfs_0)) %>% 
  select(youthidCUF, y1_bfs_1:y1_bfs_5) %>% 
  pivot_longer(cols = y1_bfs_1:y1_bfs_5,
               names_to = "best_friends",
               values_to = "friendid") %>% 
  select(youthidCUF, friendid) 

length(unique(w1_pos_long$youthidCUF))

#no_friends <- w1_pos$youthidCUF[w1_pos$y1_bfs_0 == 1]
#no_friends_id
```

```{r count by best friends}
w1_bfs <- w1_pos_long %>% 
  mutate(hasfriend = ifelse(is.na(friendid),0,1)) %>% 
  group_by(youthidCUF) %>% 
  summarise(num_bfs = sum(hasfriend)) 

descr::freq(w1_bfs$num_bfs, plot = FALSE)
```

```{r count mutual friends}
students <- unique(w1_bfs$youthidCUF)
fred_mat <- as.data.frame(students) %>% 
  mutate(n_mutual_bfs = 0)
colnames(fred_mat)[1] <- "youthidCUF"

for (i in 1 : nrow(fred_mat)) {
  id = fred_mat[i,1]
  dt = w1_pos_long %>% filter(youthidCUF == id)
  bfs = unique(dt$friendid)
  for (bf in bfs) {
    bf_data = w1_pos_long %>% filter(youthidCUF == bf)
    if (id %in% unique(bf_data$friendid)){
      fred_mat[i,2] = fred_mat[i,2] + 1
      
    } else {next}
  }
}

descr::freq(fred_mat$n_mutual_bfs)
```

```{r merge positive friends counts, eval=FALSE}
w1_pos_yc <- merge(fred_mat,w1_bfs,by="youthidCUF")
write_dta(w1_pos_yc, "Data-Stata/our_data/w1_pos_yc.dta")
```

## Friendship networks in all schools

```{r network in all school}
w1_pos_long_net <- w1_pos_long %>% 
  filter(!is.na(friendid)) %>% 
  mutate(youthidCUF = as.character(youthidCUF),
         friendid = as.character(friendid))

bfs_nw_all <- network(w1_pos_long_net[1:300,], matrix.type = "edgelist")
#Or we can do it by using the adjacency matrix:
#nw2 <- network(mat, directed = TRUE)
ggnet2(bfs_nw_all,
       size = 3,
       label = TRUE,
       label.size = 3,
       arrow.size = 3,
       arrow.gap = .03)
```

## Getting network measures

Plotting networks is nice, but sometimes you want to take social network measures into account for your analyses. For example, we can use social network measures as predictors of something. We will show you how to calculate a bunch of network measures and then show you how to use some in a linear regression model. We will remake our networks to be directional.

#### Construct Adjacency matrix

```{r construct adjacency matrix ALL}
students <- unique(w1_pos_long$youthidCUF)
mat <- matrix(0, nrow=length(students), ncol = length(students))
colnames(mat) <- as.character(students)
rownames(mat) <- as.character(students)
# A glimpse of what this look lik
mat[1:4,1:4]
```

#### Fill Adjacency Matrix

```{r fill adjacency matrix ALL}
i <- 1
for (i in 1:length(students)) {
  dt <- as.data.frame(w1_pos_long %>% filter(!is.na(friendid)))
  row.index <- which(students == dt[i,1])
  col.index <- which(students == dt[i,2]) # col 3 is best friends 1
  mat[row.index, col.index] <- 1
}
mat[1:5,1:5]
```

#### Construct Network for All Schools

```{r build network ALL}
bfs_nw_all <- network(mat, 
                      matrix.type = "adjacency",
                      ignore.eval=FALSE,
                      names.eval="true")
```

### Density

The function `gden()` returns how dense a network is. If there are many ties the network is called dense. If there are few ties the network is called sparse. `gden()` returns **the sum of tie values by the number of possible ties.** For binary data it thus reflects the proportion of ties of all possible ties and for valued data it reflects the average strength of all possible ties.

```{r network density}
gden(bfs_nw_all)
```

### Transitivity

Transitivity is a measure that reflects **to what extent nodes in networks are connected by shared connections.** For example, in friendship networks we often see that friends of friends are often friends too. You can get this measure with `gtrans()`

```{r transitivity}
gtrans(bfs_nw_all)
```

### Centrality

There are many measures of centrality. They can be **on the network level** or **on the node level**. A general network level indication of **how centralized a network is the Freeman measure of centrality.**

For individual nodes, you can use the `degree()` command. Same here.

```{r node centrality,eval=FALSE}
degree(bfs_nw_all, cmode = "indegree")
degree(bfs_nw_all, cmode = "outdegree")
degree(bfs_nw_all)
```
