---
title: "Covariate Exploration"
output: html_document
date: "2023-01-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## R Markdown

```{r import network data}
library(dplyr)
library(tidyverse)
library(readstata13)
library(ggplot2)
w1_netm <- read.dta13("Data-Stata/our_data/w1_netm.dta", convert.factors = F)
w1_ym <- read.dta13("Data-Stata/our_data/w1_ym.dta", convert.factors = F)
```

```{r include vars}
covars <- c("youthidCUF","classidCUF","schoolidCUF","schtype_geRV","y1_sex","y1_doby","y1_nationRV","y1_lpsc1","y1_lpsc2","y1_lpsc3","y1_lpsc4","y1_fact","y1_sat2")

main1 <- w1_ym[covars]
main1[main1 < 0] <- NA 
```

```{r}
ggplot(data=main1)+
  geom_density(aes(x=y1_pmonwRV),na.rm = TRUE)
```

```{r}
main1 <- merge(main1, w1_netm, by = "youthidCUF")

# change gender/ nationality / school type as factor
main1 <- main1 %>% 
  mutate(sex = as.factor(y1_sex),
         schooltype = as.factor(schtype_geRV),
         nationality = as.factor(y1_nationRV))
```


```{r}
library(gridExtra)
library(ggpubr)
##note these visualizaions do NOT INCLUDE COVARIATES!!! S
#we should create new visualizations after fitting models that show the relationships controlling for other covariates

main1 <- main1 |> 
  mutate(pos_isolate = as.factor(pos_isolate)) |> 
  mutate(neg_isolate = as.factor(neg_isolate))

main1 |> 
  select(y1_sat2, neg_indegree, neg_outdegree, pos_indegree, pos_outdegree) |>  
  mutate_at(vars(-c(y1_sat2)), scale) |> 
  pivot_longer(cols = -c(y1_sat2),
               names_to = "network_trait", 
               values_to = "value") |> 
  ggplot(aes(y = y1_sat2, x = value, color = network_trait)) +
  facet_wrap(~network_trait, labeller = function(variable, value){
    value <- gsub("neg_", "Negative ", value)
    value <- gsub("pos_", "Positive ", value)
    return(value)
  }) +
  stat_summary(geom = "line", fun = "mean") +
  scale_color_brewer(palette = "Paired") +
  labs(y = 'Self-Rated School Satisfaction\n(1-10)', title = 'How does student network traits in class relate to their school satisfaction?')+
  theme(legend.position = "none")



p1 <- ggplot(aes(x = pos_isolate, y = y1_sat2), data = main1)+
  stat_summary(geom = "bar", fun = "mean")+
  labs(x = "Positive Isolate Indicator", y = "Mean Satisfaction in School")

p2 <- ggplot(aes(x = neg_isolate, y = y1_sat2), data = main1)+
  stat_summary(geom = "bar", fun = "mean")+
  labs(x = "Negative Isolate Indicator", y = "Mean Satisfaction in School")

bar_plots <- ggarrange(p1, p2, ncol = 2)
bar_plots

p1box <- ggplot(aes(x = pos_isolate, y = y1_sat2), data = main1)+
  geom_boxplot()+
  labs(x = "Positive Isolate Indicator", y = "Mean Satisfaction in School")

p2box <- ggplot(aes(x = neg_isolate, y = y1_sat2), data = main1)+
  geom_boxplot()+
  labs(x = "Negative Isolate Indicator", y = "Mean Satisfaction in School")

box_plots <- ggarrange(p1box, p2box, ncol = 2)
box_plots

#"How is being an Isolate in a Network related to School Satisfaction?")

```

### Demographics & School-related Variables

#### School Type (Germany) --- `schtype_geRV`

-   1 = Lower secondary school

-   2 = School combining several tracks

-   3 = Intermediate secondary school

-   4 = Comprehensive school / Rudolf-Steiner School

-   5 = Upper secondary school

-   6 = School for special needs

#### Gender --- `y1_sex`

-   1 = Boy

-   2 = Girl

#### Age --- `y1_doby` (when were you born? Year)

#### Migration age --- `y1_migage` (how old were you when you moved to [survey country]?)

#### Nationality --- `y1_nationRV`

-   1 = Only survey country nationality

-   2 = Survey country and other nationality

-   3 = Only other nationality

#### Language

##### `y1_lpsc1` How well can you speak German [survey country language] ?

-   1 = Not at all

-   2 = Not well

-   3 = Well

-   4 = Very well

-   5 = Excellently

##### `y1_lpsc2` How well can you understand German?

##### `y1_lpsc3` read German \| `y1_lpsc4` write German

#### Problem behavior / Deviance & Delinquency at school

##### `y1_pbsch1` How often do you argue with teacher?

-   1 = every day

-   2 = once or several times a week

-   3 = once or several times a month

-   4 = less often

-   5 = never

##### `y1_pbsch2` --- how often do you get a punishment in school (e.g. being kept in detention, sent out of class, writing lines)?

##### `y1_pbsch3` --- how often do you skip a lesson without permission?

##### `y1_pbsch4` --- how often do you come late to school?

##### `y1_del1 : y1_del4` --- Delinquent behavior I : IV

### Family

#### Family Relation

##### `y1_contm` --- how often do you usually see your mother?

##### `y1_gawm` --- how well do you get along with your mother?

##### `y1_contf` --- how often do you usually see your father?

##### `y1_gawf` --- how well do you get along with your father?

#### Economic Situation

##### `y1_pmon| y1_pmonwRV | y1_pmonmRV` --- Pocket money (general / weekly / monthly amount)

`y1_pmon` -- Do you get pocket money from your parents? - 1 = yes, each week - 2 = yes, each month - 3 =

##### `y1_fact` --- can't afford friends' activities

### Add Some Demo covariates to our model
```{r}

w1_predictors_clean <- read.dta13("Data-Stata/our_data/w1_predictors_clean.dta", convert.factors = F)
main2 <- w1_ym[covars]
main2[main2 < 0] <- NA
main2 <- merge(main2, w1_predictors_clean)
summary(lm(satisfaction ~ pos_indegree + pos_outdegree + pos_isolate + neg_indegree + neg_outdegree + neg_isolate + n_mutual_bfs + as.factor(classidCUF) +  as.factor(y1_nationRV) + as.factor(schtype_geRV) + as.factor(y1_sex), data=main2))
```
