---
title: "EDA & Data Manipulation -- More Covariates"
output: html_document
date: "2023-01-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Exploratory Data Analysis and More Data Manipulation

```{r import network data}
library(dplyr)
library(tidyverse)
library(readstata13)
library(ggplot2)
w1_netm <- read.dta13("Data-Stata/our_data/w1_netm.dta", convert.factors = F)
w1_ym <- read.dta13("Data-Stata/our_data/w1_ym.dta", convert.factors = F)
```

```{r include vars}
covars <- c("youthidCUF","classidCUF","schoolidCUF","schtype_geRV","y1_sex","y1_doby","y1_nationRV","y1_lpsc1","y1_lpsc2","y1_lpsc3","y1_lpsc4","y1_fact","y1_sat2","y1_pbsch1","y1_pbsch2","y1_pbsch3","y1_pbsch4","y1_del1","y1_del2","y1_del3","y1_del4","y1_vict1","y1_vict2","y1_vict3","y1_pdisc1","y1_pdisc2","y1_pdisc3","y1_pdisc4","y1_date","y1_contm","y1_gawm","y1_contf","y1_gawf","y1_penc1","y1_penc2","y1_penc3","y1_fcoh1","y1_fcoh2","y1_fcoh3","y1_fcoh4","y1_fcoh5","y1_pdem1","y1_pdem2","y1_pdem3","y1_pdem4","y1_pdem5","y1_pdem6","y1_pdem7","y1_pinv1","y1_pinv2","y1_pinv3")

# deviance and delinquency in school -- y1_pbsch1 : y1_del4
# perceived victim / discrimination in school -- y1_vict1 : y1_pdisc4
# romantic relationship -- y1_date (currently dating?)
# family relations -- y1_contm : y1_contf
# parental encouragement -- y1_penc1 : y1_penc3
# family cohesion -- y1_fcoh1 : y1_fcoh5
# parenting style -- demandingness -- y1_pdem1:y1_pdem7
# parenting style -- involvement -- y1_pinv1:y1_pinv3

# well-being
# self esteem -- y1_sest1:y1_sest4

# attitudes toward school
# teachers encouragement y1_tenc1 : y1_tenc3

main1 <- w1_ym[covars]
main1 <- w1_ym %>% 
  select(youthidCUF, classidCUF, schoolidCUF, schtype_geRV,
         y1_sex, y1_doby, y1_nationRV, y1_lpsc1:y1_lpsc4, 
         y1_fact, y1_sat2, y1_sat1, y1_pbsch1:y1_pbsch4, 
         y1_del1:y1_del4, y1_vict1:y1_vict3,y1_pdisc1:y1_pdisc4,
         y1_date, y1_contm,y1_gawm, y1_contf, y1_gawf, 
         y1_penc1:y1_penc3, y1_fcoh1:y1_fcoh5, y1_pdem1:y1_pdem7,
         y1_pinv1:y1_pinv3, y1_sest1:y1_sest4,
         y1_tenc1:y1_tenc3
         )

main1[main1 < 0] <- NA 
main1$age <- 2010 - main1$y1_doby
main1 <- merge(main1, w1_netm, by = c("youthidCUF","classidCUF"))

write_csv(main1,"Data-Stata/our_data/all_data.csv")
```

#### School Type (Germany) --- `schtype_geRV`

-   1 = Lower secondary school

-   2 = School combining several tracks

-   3 = Intermediate secondary school

-   4 = Comprehensive school / Rudolf-Steiner School

-   5 = Upper secondary school

-   6 = School for special needs

```{r school type}
# univariate analysis of school type
descr::freq(main1$schtype_geRV)
```

## Cultural Integration
### Language measures -- proficiency in survey country language (German)
-   5-point Likert type (1 = Not at all, 2 = Not well ... 5 = Excellently)
-   four sub-measures regarding speak, read, write, understand

Data manipulation --> recode them into 0 to 4 for each measure and sum them up
```{r language}
# recode multiple variables whose levels are 1 to 5 into 0 to 4
main2 <- main1 %>% 
  mutate_at(vars(y1_lpsc1:y1_lpsc4), funs(.-1))

# check summary stats 
print("Before recoding")
summary((main1 %>% select(y1_lpsc1:y1_lpsc4)))
print("After recoding")
summary((main2 %>% select(y1_lpsc1:y1_lpsc4)))

main2 <- main2 %>% 
  mutate(language = y1_lpsc1 + y1_lpsc2 + y1_lpsc3 + y1_lpsc4)
descr::freq(main2$language)
```

## Socio-Demography and Migration
### Gender --- `y1_sex`

-   1 = Boy

-   2 = Girl

```{r gender type}
# add factor variable of gender
main2$gender = ifelse(main2$y1_sex == 1, "Boy", "Girl")
descr::freq(main2$gender, plot = FALSE)
```

### Age --- `y1_doby` (when were you born? Year)
```{r age}
descr::freq(main2$age, plot = FALSE)
```

### Nationality --- `y1_nationRV` 

-   1 = Only survey country nationality

-   2 = Survey country and other nationality

-   3 = Only other nationality

```{r nationality}
main2 <- main2 %>% 
  mutate(nationality = case_when(
    y1_nationRV == 1 ~ "German only",
    y1_nationRV == 2 ~ "German and other nationality",
    y1_nationRV == 3 ~ "Only other nationality",),
    nationality = factor(nationality, 
                            levels=c("German only",
                                     "German and other nationality",
                                     "Only other nationality")))

# check for recoding
print("Before recoding")
descr::freq(main1$y1_nationRV,plot = FALSE)
print("After recoding")
descr::freq(main2$nationality,plot = FALSE)
```

##### Correlation Matrix

```{r corr,eval=FALSE}
library(corrplot)
library(Hmisc)

mydata.cor = cor(na.omit(main1[,-c(1:3,6)]), method = c("spearman"))
mydata.cor
corrplot(mydata.cor, method = "color")
```

##### Summary Statistics
```{r, eval=FALSE}
# change gender/ nationality / school type as factor
summary(main1 %>% 
  mutate(classidCUF = as.factor(classidCUF),
         schoolidCUF = as.factor(schoolidCUF),
         y1_sex = as.factor(y1_sex),
         schtype_geRV = as.factor(schtype_geRV),
         y1_nationRV = as.factor(y1_nationRV),
         y1_fact = as.factor(y1_fact)) %>% 
  select(-youthidCUF))
```

## Economic situation
### `y1_fact` --- can't afford friends' activities
-   1 = Always
-   2 = Often
-   3 = Sometimes
-   4 = Never

Note that high values in this variable represent low economic situation
```{r eco}
descr::freq(main1$y1_fact, plot = FALSE)
```

## Structual Integration
### Deviance & Delinquency at school
#### `y1_pbsch1` How often do exihibit following problem behaviors in school?

argue with teacher | get a punishment in school (e.g. being kept in detention, sent out of class, writing lines) | skip a lesson without permission? | come late to school?

Response:
-   1 = every day

-   2 = once or several times a week

-   3 = once or several times a month

-   4 = less often

-   5 = never
```{r problem behaviors}

# recoding problem behavioral measures
main2 <- main2 %>% 
  mutate_at(vars(y1_pbsch1:y1_pbsch4), ~ 5-.) %>% 
  mutate(problem_behaviors = y1_pbsch1 + y1_pbsch2 + y1_pbsch3 + y1_pbsch4)
# check before recoding
print("Before recoding")
descr::freq(main1$y1_pbsch1, plot = FALSE)
print("After recoding")
descr::freq(main2$y1_pbsch1, plot = FALSE)
print("Frequency of problem behaviors")
descr::freq(main2$problem_behaviors)
```

#### `y1_del1` --- Delinquent behavior: Deliberately damaged things that were not yours
-   1 = yes
-   2 = no
#### `y1_del2` --- Delinquent behavior: Stolen something from a shop/from someone else?
#### `y1_del3` --- Delinquent behavior: Carried a knife or weapon
#### `y1_del1` --- Delinquent behavior: Been very drunk
```{r delinquency behavior}
main2 <- main2 %>% 
  mutate_at(vars(y1_del1:y1_del4), 
            funs(replace(., . == 2,0)))

print("Before recoding")
summary(main1 %>% mutate_at(vars(y1_del1:y1_del4), as.factor) %>% select(y1_del1 : y1_del4))

print("After recoding")
summary(main2 %>% mutate_at(vars(y1_del1:y1_del4), as.factor) %>% select(y1_del1 : y1_del4))

# if or not they have delinquency behavior
main2 <- main2 %>% 
  mutate(delinquency_history = ifelse(rowSums(
    select(.,y1_del1:y1_del4) == 1) > 0,
    'yes','no'),
    delinquency_count = y1_del1 + y1_del2 + y1_del3 + y1_del4)

print("Delinquency history frequency")
descr::freq(main2$delinquency_history,plot = FALSE)
print("Delinquency count frequency")
descr::freq(main2$delinquency_count,plot = FALSE)
```

### Discrimination at school
Note that we donnot use perceived discrimination measures as they occur outside school setting!
#### `y1_vict1:y1_vict3` -- Victim at school following types:
scared of other students | teased by other students | bullied by other students 

Responses:
-   1 = every day
-   2 = once or several times a week
-   3 = less often
-   4 = never
```{r victim}
main2 <- main2 %>%
  mutate_at(vars(y1_vict1:y1_vict3),  ~4-.)

print("Before recoding")
summary(main1 %>% mutate_at(vars(y1_vict1:y1_vict3), as.factor) %>% select(y1_vict1:y1_vict3))
print("After recoding")
summary(main2 %>% mutate_at(vars(y1_vict1:y1_vict3), as.factor) %>% select(y1_vict1:y1_vict3))

main2 <- main2 %>% 
  mutate(victim_history = ifelse(rowSums(
    select(.,y1_vict1:y1_vict3) == 1) > 0,
    1,0)) %>% 
  mutate_at(vars(y1_vict1:y1_vict3), 
         funs(ifelse(. == 3, 5, # every day
                     ifelse(. == 2, 4, # once or several times a week
                            ifelse(. == 1, 2, .) # less often
                            )
                     )
              )
         ) # never
main2 <- main2 %>% 
  mutate(victim_history_score = y1_vict1 + y1_vict2 + y1_vict3)
descr::freq(main2$victim_history,plot = FALSE)
descr::freq(main2$victim_history_score)
print("After recoding stage two")
summary(main2 %>% mutate_at(vars(y1_vict1:y1_vict3), as.factor) %>% select(y1_vict1:y1_vict3))
```


### Family 

#### Family Relation
##### `y1_gawm | y1_gawf` --- how well do you get along with your mother/father?
-   1 = every very well
-   2 = well
-   3 = not so well
-   4 = not well at all
```{r parents relation}
# if or not they have delinquency behavior
main2 <- main2 %>% 
  mutate_at(vars(y1_gawm,y1_gawf), ~4-.) %>% 
  mutate(parent_relation = round((y1_gawf + y1_gawm)/2)) # average score of how well they get along with parents

print("Before recoding")
summary(main1 %>% mutate_at(vars(y1_gawm,y1_gawf), as.factor) %>% select(y1_gawm,y1_gawf))

print("After recoding")
summary(main2 %>% mutate_at(vars(y1_gawm,y1_gawf), as.factor) %>% select(y1_gawm,y1_gawf))

print("Parental relation")
descr::freq(main2$parent_relation,plot = FALSE)
```


## School satisfication
```{r outcome}
descr::freq(main2$y1_sat2)
```

### Self-esteem
#### `y1_sest1:y1_sest4` following measures:
I have a lot of good qualities | have a lot to be proud of | like myself just the way I am | think things will go well for me in the future 

Response:
-   1 = strongly agree
-   2 = agree
-   3 = neither agree nor disagree
-   4 = disagree
-   5 = strongly disagree

```{r save as json file, eval=FALSE}
jsonlite::write_json(main2, "Data-Stata/our_data/CILS4EU.json")
write_csv(main2, "Data-Stata/our_data/all_main.csv")
```

