---
title: "friends_network"
output: html_document
date: "2023-01-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```
```{r}
library(dplyr)
library(readstata13)
library(tidyverse)
library(sna)
library(GGally)
```

## Positive Friendship Network
```{r data preprocessing}
w1_yc <- read.dta13("Data-Stata/our_data/w1_yc.dta",convert.factors =F)
class1 <- w1_yc %>% 
  select(-y1_intdat_ycRV, -y1_vers_yc)
# look at the structur of data
#str(w1_class)
```

In Sanne's code: 
"# add health students mention on average 6 or 7 friends. 
 # but these are american students who have lower thresholds fo friendship nominations
 # drop cases according to Kruse & Jacobs
Q: do we also want to drop these cases?
```{r drop cases according to Kruse & Jacobs}
# add health students mention on average 6 or 7 friends. 
# but these are american students who have lower thresholds for friendship nominations
# drop cases according to Kruse & Jacobs
class1$n <- 1
myvars <- c("n", "y1_nnomineeG")
npupils <- aggregate(class1[myvars], sum, by = list(class1$classidCUF))
keep <- npupils[npupils$n > 10 & npupils$y1_nnomineeG <=4, 1]

class1 <- class1[class1$classid %in% keep, ]
class1 <- class1[class1$y1_classid_matchG == 1, ]
class1 <- class1[class1$y1_p_participantsG >= .75, ]
class1 <- class1[class1$y1_p_invalidnomsG <= .1, ]
```

```{r select positive friendship measures}
# code out missing values
# select friendship measures and set missing values to NA
pos_vars <- c("y1_bfs_1", "y1_bfs_2", "y1_bfs_3", "y1_bfs_4", "y1_bfs_5")
class1[pos_vars][class1[pos_vars] < 0] <- NA
descr::freq(class1$y1_bfs_0)
```

Q: For those who did not report having best friend in class, do we remove them or include them and make their # class best friends = 0?
```{r reshape to long}
# reshape to long
class1.l <- class1 %>% 
  select(youthidCUF, y1_bfs_1:y1_bfs_5) %>% 
  pivot_longer(cols = y1_bfs_1:y1_bfs_5,
               names_to = "best_friends",
               values_to = "alteridCUF") %>% 
  select(youthidCUF, alteridCUF) %>% 
  na.omit() # remove NAs for building network

### Sanne's code
#myvars <- c("youthidCUF", "y1_bfs_1", "y1_bfs_2", "y1_bfs_3", "y1_bfs_4", "y1_bfs_5")
#class1.l <- reshape(class1[myvars], varying = myvars[-1], v.names = "alterid", timevar = "friend", idvar = "youthidCUF", times =myvars[-1], direction = "long")
```

```{r data preprocess for network}
myids <- as.data.frame(unique(class1[c("youthidCUF","classidCUF")]))
classids <- unique(class1$classidCUF)
class1.l <- merge(class1.l, myids, by = "youthidCUF", all.x = TRUE)
classids <- unique(class1.l$classid)

nclass <- length(unique(class1.l$classid))
nclass

# make sure you take only the last two digits of youthid otherwise sna
# thinks that there are a thousand actors
class1.l <- class1.l %>% 
  mutate(youthid = youthidCUF %% 100,
         alterid = alteridCUF %% 100) %>% 
  select(youthid, alterid, classidCUF)
```

```{r make network for each class}
el.list <- list()
class.list <- list()

for (i in 1:nclass){
  el.list[[i]] <- class1.l[class1.l$classidCUF == classids[i],]
  class.list[[i]] <- classids[i]
}

net.list <- lapply(el.list, as.network.matrix, matrix.type = "edgelist", 
 ignore.eval = FALSE, names.eval = "classidCUF")	
```

### Visualization of network 1 -- classidCUF 102001
Q:
```{r network viz}
ggnet2(net.list[[1]],
       label = TRUE,
       label.size = 3,
       arrow.size = 3,
       arrow.gap = .03)
```

### Calculate positive friendship network measures 
```{r network measures}
mat.list <- lapply(net.list, as.matrix)
indegree.list <- lapply(mat.list, degree, cmode = "indegree", gmode = "digraph", diag = FALSE)
outdegree.list <- lapply(mat.list, degree, cmode = "outdegree", gmode = "digraph", diag = FALSE)
#bonpow.list <- lapply(mat.list, bonpow, diag = FALSE)
  # error? singular?
```
```{r unlist network measures}
a.list <- lapply(net.list, get.vertex.attribute, "vertex.names")
netm <- mapply(cbind, a.list, class.list, indegree.list, outdegree.list)
netm <- lapply(netm, as.data.frame)
netm <- as.data.frame(do.call(rbind, netm))
names(netm) <- c("youthid", "classid", "indegree", "outdegree")

# computate isolate measure
netm <- netm %>% 
  mutate(isolate = ifelse(indegree == 0 & outdegree == 0, 1, 0),
         youthid_chr = as.character(youthid),
         classid_chr = as.character(classid),
         youthid_nchr = nchar(youthid_chr) # count the number of characters
         )

# add 0 to single digit youthid
netm$youthid_chr[netm$youthid_nchr == 1] <- paste0("0", netm$youthid_chr[netm$youthid_nchr == 1])

# change to numeric youthid
netm$youthid <- as.numeric(paste0(netm$classid_chr, netm$youthid_chr))
```

```{r save positive friendship network, eval=FALSE}
haven::write_dta(netm, "Data-Stata/our_data/w1_pos_netm.dta")
```

