---
title: "Data Cleaning"
output: html_document
date: "2023-01-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r load packages}
library(dplyr)
library(ggplot2)
library(descr) # for using freq() funcion
```
## Tracking data Wave 1
```{r wave1 youth classmates}
freq(w123_tracking$ps_w1_yc, plot = FALSE)
```
## Tracking data - Wave 2

```{r wave2 youth}
freq(w123_tracking$ps_w2_yRV, plot = FALSE)
```

```{r wave2 youth main}
freq(w123_tracking$ps_w2_ym, plot = FALSE)
```

### Wave 2 DOES NOT have friends questionnaire!

### 738 students DID NOT respond to Wave 2 classmates questionnaire.

```{r wave2 youth classmates}
freq(w123_tracking$ps_w2_yc, plot = FALSE)
```

## Tracking data - Wave 3

```{r wave3 youth}
freq(w123_tracking$ps_w3_yRV, plot = FALSE)
```

### All 2498 participants responded to Wave 3 Youth Main questionnaire.

```{r wave3 youth main}
freq(w123_tracking$ps_w3_ym, plot = FALSE)
```

### 17 participants DID NOT respond to Wave 3 Youth Friends (5 best friends) questionnaire.

```{r wave3 youth friend}
freq(w123_tracking$ps_w3_yf, plot = FALSE)
```

### Wave 3 DOES NOT have classmates questionnaire!

## Data Wrangling First Stage --> W1 & W2 

We decided to (preliminarily) work with the wave 1 and wave 2 data as wave 3 does not contain classmates component. And the classmate data contains richer information about weak/strong tie strengths of friendships as well as unique identifiers of each student that allow us to connect each individual to form the friendship network.

### 1. Excluding students who did not respond to the wave 1 & 2 classmates questionnaire.
```{r excluding respondents from w2, eval=FALSE}
# list of student ids needed to be excluded from cross-sectional analyses of wave 1 and wave 2 data
excluded_student_ids <- w123_tracking %>% 
  filter(ps_w2_yc == 0 | ps_w1_yc == 0) %>% 
  select(youthidCUF)

length(excluded_student_ids$youthidCUF)
# 9 did not respond to wave 1 classmates
# 738 did not repond to wave 2 classmates

w1_yc_updated <- w1_classmate %>% 
  filter(!youthidCUF %in% excluded_student_ids$youthidCUF)
write_dta(w1_yc_updated, "Data-Stata/our_data/w1_yc.dta")

w1_ym_updated <- w1_main %>% 
  filter(!youthidCUF %in% excluded_student_ids$youthidCUF)
write_dta(w1_ym_updated, "Data-Stata/our_data/w1_ym.dta")

w1_yf_updated <- w1_friend %>% 
  filter(!youthidCUF %in% excluded_student_ids$youthidCUF)
write_dta(w1_yf_updated, "Data-Stata/our_data/w1_yf.dta")

w2_yc_updated <- w2_classmate %>% 
  filter(!youthidCUF %in% excluded_student_ids$youthidCUF)
write_dta(w2_yc_updated, "Data-Stata/our_data/w2_yc.dta")

w2_ym_updated <- w2_main %>% 
  filter(!youthidCUF %in% excluded_student_ids$youthidCUF)
write_dta(w2_ym_updated, "Data-Stata/our_data/w2_ym.dta")
```


## W1 friends 
Variable: y1_schf1 / y1_schf2 / y1_schf3, etc.
Does he/she go to your school? Friend1
1 - Yes, same class
2 - Yes, but different class
3 - No, goes to another school
4 - No. has finished schooling
```{r w1 friends}
freq(w1_friend$y1_schf1, plot = FALSE)
freq(w1_friend$y1_schf2, plot = FALSE)
freq(w1_friend$y1_schf3, plot = FALSE)
freq(w1_friend$y1_schf4, plot = FALSE)
freq(w1_friend$y1_schf5, plot = FALSE)

```

## W1 Classmates
```{r w1 friends}
unique(w1_classmate$y1_bfs_0)
freq(w1_classmate$y1_bfs_0)
unique(w1_friend$y1_f1G)
```

